*Prochlorococcus marinus* responses to light and oxygen

### 2024-02-26 Code/Manuscript_O2.Rmd
Add OPP citation

Low-Light clades thrive in deeper ocean waters, extending beyond 200 m in depth [@partenskyProchlorococcusMarinePhotosynthetic1999], where only ~1% of the surface irradiance penetrates, primarily in the blue (450 nm) to green (520 nm) spectral range [@holtropVibrationalModesWater2021]. Clade LLIV, including cultured strain MIT9313, falls near the base of the *Prochlorococcus* radiation, and is characterized by preference for low light, typical of depths from 120 m to 200 m [@partenskyProchlorococcusMarinePhotosynthetic1999]. LLIV members are, as yet, the only cultured strains to have been found in Oxygen Minimum Zones (OMZ).  Some as yet uncultured *P. marinus* strains in clades LLV and LLVI also thrive in OMZ of the subtropical Atlantic and Pacific Oceans, where dissolved oxygen concentrations [O~2~] can be less than 20 µM [@ulloaCyanobacteriumProchlorococcusHas2021; @lavinNovelLineagesProchlorococcus2010; @goerickeNovelNicheProchlorococcus2000; @garcia-robledoCrypticOxygenCycling2017; @johnsonEnergeticsGrowthKinetics1999]. *Prochlorococcus marinus* LL ecotypes may indeed dominate the phytoplankton within OMZ
[@lavinNovelLineagesProchlorococcus2010;
@ulloaCyanobacteriumProchlorococcusHas2021; @partenskyProchlorococcusAdvantagesLimits2010], and may be net O~2~ consumers [@partenskyComparisonPhotosyntheticPerformances2018].

Clades LLII and LLIII are grouped together as the second oldest phylogenetic lineages diversifying in the *P. marinus* radiation, and share an affinity for low light; LLIII cultured strain SS120, was isolated from 135 m depth in the Sargasso Sea.

Clade LLI  includes cultured strains NATL1A and NATL2A, which prefer moderate irradiances typical between 30 and 100 m depth.

@@

We used the OceansMap Protein Portal (OPP; https://proteinportal.whoi.edu/) to analyze the distribution of proteins from clades of *P. marinus* in samples tanken across a range of [O~2~] and depth, which in turn correlates to peak light at the site of sampling. We therefore analyzed the growth and physiological responses of representative strains from three clades of *P. marinus* under a matrix of [O~2~], light levels, spectral waveband ranges, and photoperiods, to approximate eco-physiological conditions representative of current and hypothetical future ocean zones. *P. marinus* MED4, a clade HLI strain, was isolated near the ocean surface (5m depth) of the Mediterranean Sea where [O~2~] is near saturation, light levels are high and spectral bias from full solar irradiance is minimal. *P. marinus* SS120, a clade LLIII strain, was isolated from the Sargasso Sea at a depth of 120 m, while *P. marinus* MIT9313, a clade LLIV strain, was isolated from the North Atlantic Gulf Stream at a depth of 135 m. At these depths, light attenuation and spectral shifts occur, resulting in low blue and green light, while [O~2~] varies from near-surface saturation levels to decreased concentrations, but does not necessarily decrease systematically with depth [@billheimerOxygenSeasonalityUtilization2021].

@@

The OceansMap Protein Portal (OPP; https://proteinportal.whoi.edu/) is an open access online data repository (Woods Hole Oceanographic Institute, WHOI) of mass spectroscopy data on marine microbial peptides, sampled from various depths and locations worldwide. We screened a subset of the OPP for proteins annotated as from *Prochlorococcus* strains, to identify differential strategies employed by strains living at varying depths and oxygen levels within the marine water column.  We focussed on proteins mediating photosynthesis and protein metabolism. The samples for metaproteomic analyses were collected from 12 locations in the tropical North Pacific ocean along 150 W from 18 N of the equator between October 1, 2011 and October 25, 2011 during the voyage of the R/V Kilo Moana MetZyme cruise KM1128 (https://www.rvdata.us/search/cruise/KM1128; original datasets in the Biological and Chemical Oceanography Data Management Office repository; https://www.bco-dmo.org/project/2236).

>>>>

Low-Light clades thrive in deeper ocean waters, extending beyond 200 m in depth [@partenskyProchlorococcusMarinePhotosynthetic1999], where only ~1% of the surface irradiance penetrates, primarily in the blue (450 nm) to green (520 nm) spectral range [@holtropVibrationalModesWater2021]. Clade LLIV, including cultured strain MIT9313, falls near the base of the *Prochlorococcus* radiation, and is characterized by preference for low light, typical of depths from 120 m to 200 m [@partenskyProchlorococcusMarinePhotosynthetic1999]. LLIV members are, as yet, the only cultured strains to have been found in Oxygen Minimum Zones (OMZ).  Some as yet uncultured *P. marinus* strains in clades LLV and LLVI also thrive in OMZ of the subtropical Atlantic and Pacific Oceans, where dissolved oxygen concentrations [O~2~] can be less than 20 µM [@ulloaCyanobacteriumProchlorococcusHas2021; @lavinNovelLineagesProchlorococcus2010; @goerickeNovelNicheProchlorococcus2000; @garcia-robledoCrypticOxygenCycling2017; @johnsonEnergeticsGrowthKinetics1999]. *Prochlorococcus marinus* LL ecotypes may indeed dominate the phytoplankton within OMZ
[@lavinNovelLineagesProchlorococcus2010;
@ulloaCyanobacteriumProchlorococcusHas2021; @partenskyProchlorococcusAdvantagesLimits2010], and may be net O~2~ consumers [@partenskyComparisonPhotosyntheticPerformances2018]. Clades LLII and LLIII are grouped together as the second oldest phylogenetic lineages diversifying in the *P. marinus* radiation, and share an affinity for low light; LLIII cultured strain SS120, was isolated from 135 m depth in the Sargasso Sea. Clade LLI  includes cultured strain NATL2A, which prefers moderate irradiance typically between 30 and 100 m depth.

@@

We used the OceansMap Protein Portal (OPP; https://proteinportal.whoi.edu/) [@saitoDevelopmentOceanProtein2021] to analyze the distribution of proteins from clades of *P. marinus* in samples tanken across a range of [O~2~] and depth, which in turn correlates to peak light at the site of sampling. We therefore analyzed the growth and physiological responses of representative strains from three clades of *P. marinus* under a matrix of [O~2~], light levels, spectral waveband ranges, and photoperiods, to approximate eco-physiological conditions representative of current and hypothetical future ocean zones. *P. marinus* MED4, a clade HLI strain, was isolated near the ocean surface (5m depth) of the Mediterranean Sea where [O~2~] is near saturation, light levels are high and spectral bias from full solar irradiance is minimal. *P. marinus* SS120, a clade LLIII strain, was isolated from the Sargasso Sea at a depth of 120 m, while *P. marinus* MIT9313, a clade LLIV strain, was isolated from the North Atlantic Gulf Stream at a depth of 135 m. At these depths, light attenuation and spectral shifts occur, resulting in low blue and green light, while [O~2~] varies from near-surface saturation levels to decreased concentrations, but does not necessarily decrease systematically with depth [@billheimerOxygenSeasonalityUtilization2021].

@@

The OceansMap Protein Portal is an open access online data repository (Woods Hole Oceanographic Institute, WHOI) of mass spectroscopy data on marine microbial peptides, sampled from various depths and locations worldwide [@saitoDevelopmentOceanProtein2021]. We screened a subset of the OPP for proteins annotated as from *Prochlorococcus* strains, to identify differential strategies employed by strains living at varying depths and oxygen levels within the marine water column.  We focussed on proteins mediating photosynthesis and protein metabolism. The samples for metaproteomic analyses were collected from 12 locations in the tropical North Pacific ocean along 150 W from 18 N of the equator between October 1, 2011 and October 25, 2011 during the voyage of the R/V Kilo Moana MetZyme cruise KM1128 (https://www.rvdata.us/search/cruise/KM1128; original datasets in the Biological and Chemical Oceanography Data Management Office repository; https://www.bco-dmo.org/project/2236).
